### 2011-01-05 proof.tex
Annotated proof of output_input_at_update_times
Includes new technology for annotating proofs.
Collects proof states along with string descriptions using
side-effecting tactics that don't change the proof state.
The presentation side doesn't work too well at the moment, mostly
because LaTeX and the pretty-printer don't get along...

Also reorder simplifiedTheory so everything before
output_input_at_update_times is a dependency of it.

\documentclass{article}
\usepackage{holtexbasic}
\usepackage{alltt}
\usepackage{fixltx2e}
\begin{document}
\input{output-input-at-update-times.tex}
\end{document}