### 2011-01-05 proof.tex
Use smaller margins and no paragraph indents

\documentclass{article}
\usepackage{holtexbasic}
\usepackage{alltt}
\usepackage[a4paper,margin=2cm]{geometry}
\usepackage{parskip}
\begin{document}
\input{output-input-at-update-times.tex}
\end{document}