### 2011-01-05 proof.tex
Start annotating proofs for more theorems
Add a write_proof function that writes the last proof annotated, adding
extra rules to the grammars beforehand and removing them afterwards.

\documentclass{article}
\usepackage{holtexbasic}
\usepackage{alltt}
\usepackage[a4paper,margin=2cm]{geometry}
\usepackage{parskip}
\begin{document}
\input{Output-First.tex}
\input{Output-Source.tex}
\input{Output-Input.tex}
\end{document}