### 2011-01-05 proof.tex
Make the proof tex files end with Proof.tex and gitignore them

\documentclass{article}
\usepackage{holtexbasic}
\usepackage{alltt}
\usepackage[a4paper,margin=2cm]{geometry}
\usepackage{parskip}
\begin{document}
\input{OutputFirstProof.tex}
\input{OutputSourceProof.tex}
\input{OutputInputProof.tex}
\end{document}